---
title: "MSBI 33200 - Manual Logistic Regression"
output:
  html_document:
    df_print: paged
---

# Preamble 

Won, VERY nice work at getting this done manually! :) What I'm doing here is cleaning up your code and showing, perhaps, a more efficient way to do it. The less code you write, the lower the odds of a mistake :) 

```{r setup}
library(tidyverse) # Tidyverse loads dplyr, ggplot2, readr for you automatically
# You don't need to manually load them, although you can if you wish to be explicit
library(data.table) # I like this package for its fread() function
library(mltools) # This will help us with the one-hot encoding we need later 

diabetes <- 
  fread("supporting_files/diabetes_data_upload.csv") %>%
  rename_with(toupper) # This standardizes the case of the variables to uppercase 
```

We can accomplish all of this... 

```{{r}}
#Data Wrangling
spec(diabetes_data)
class(diabetes_data)
##Gender
unique(diabetes_data$Gender)
diabetes_data$Gender <- ifelse(diabetes_data$Gender=="Male",1,0)
diabetes_data$Gender
##Polyuria
unique(diabetes_data$Polyuria)
diabetes_data$Polyuria <- ifelse(diabetes_data$Polyuria=="Yes",1,0)
diabetes_data$Polyuria
##Polydipsia
unique(diabetes_data$Polydipsia)
diabetes_data$Polydipsia <- ifelse(diabetes_data$Polydipsia=="Yes",1,0)
diabetes_data$Polydipsia
##'sudden weight loss'
unique(diabetes_data$`sudden weight loss`)
diabetes_data$`sudden weight loss` <- ifelse(diabetes_data$`sudden weight loss`=="Yes",1,0)
diabetes_data$`sudden weight loss`
##weakness
unique(diabetes_data$weakness)
diabetes_data$weakness <- ifelse(diabetes_data$weakness=="Yes",1,0)
diabetes_data$weakness
##Polyphagia
unique(diabetes_data$Polyphagia)
diabetes_data$Polyphagia <- ifelse(diabetes_data$Polyphagia=="Yes",1,0)
diabetes_data$Polyphagia
##'Genital thrush'
unique(diabetes_data$`Genital thrush`)
diabetes_data$`Genital thrush` <- ifelse(diabetes_data$`Genital thrush`=="Yes",1,0)
diabetes_data$`Genital thrush`
##'visual blurring'
unique(diabetes_data$`visual blurring`)
diabetes_data$`visual blurring` <- ifelse(diabetes_data$`visual blurring`=="Yes",1,0)
diabetes_data$`visual blurring`
##Itching
unique(diabetes_data$Itching)
diabetes_data$Itching <- ifelse(diabetes_data$Itching=="Yes",1,0)
diabetes_data$Itching
##Irritability
unique(diabetes_data$Irritability)
diabetes_data$Irritability <- ifelse(diabetes_data$Irritability=="Yes",1,0)
unique(diabetes_data$Irritability)
##'delayed healing'
unique(diabetes_data$`delayed healing`)
diabetes_data$`delayed healing` <- ifelse(diabetes_data$`delayed healing`=="Yes",1,0)
unique(diabetes_data$`delayed healing`)
##'partial paresis'
unique(diabetes_data$`partial paresis`)
diabetes_data$`partial paresis` <- ifelse(diabetes_data$`partial paresis`=="Yes",1,0)
unique(diabetes_data$`partial paresis`)
##'muscle stiffness'
unique(diabetes_data$`muscle stiffness`)
diabetes_data$`muscle stiffness` <- ifelse(diabetes_data$`muscle stiffness`=="Yes",1,0)
unique(diabetes_data$`muscle stiffness`)
##Alopecia
unique(diabetes_data$Alopecia)
diabetes_data$Alopecia <- ifelse(diabetes_data$Alopecia=="Yes",1,0)
unique(diabetes_data$Alopecia)
##Obesity
unique(diabetes_data$Obesity)
diabetes_data$Obesity <- ifelse(diabetes_data$Obesity=="Yes",1,0)
unique(diabetes_data$Obesity)
##class
unique(diabetes_data$class)
diabetes_data$class <- ifelse(diabetes_data$class=="Positive",1,0)
unique(diabetes_data$class)
```

With this: 

```{r data-wrangling}
diabetes <- 
diabetes %>%
  mutate(across(where(is.character), as.factor))

head(diabetes,10)
```
Factors are internally represented as numbers, but are displayed to us as text. 

```{r}
summary(diabetes)
```

```{{r}}
#Checking Sample for Imbalance
xtabs(~class + Gender, data = diabetes_data)
xtabs(~class + Polyuria, data = diabetes_data)
xtabs(~class + Polydipsia, data = diabetes_data)
xtabs(~class + diabetes_data$`sudden weight loss`, data = diabetes_data)
xtabs(~class + weakness, data = diabetes_data)
xtabs(~class + Polyphagia, data = diabetes_data)
xtabs(~class + diabetes_data$`Genital thrush`, data = diabetes_data)
xtabs(~class + diabetes_data$`visual blurring`, data = diabetes_data)
xtabs(~class + Itching, data = diabetes_data)
xtabs(~class + Irritability, data = diabetes_data)
xtabs(~class + diabetes_data$`delayed healing`, data = diabetes_data)
xtabs(~class + diabetes_data$`partial paresis`, data = diabetes_data)
xtabs(~class + diabetes_data$`muscle stiffness`, data = diabetes_data)
xtabs(~class + Alopecia, data = diabetes_data)
xtabs(~class + Obesity, data = diabetes_data)
```

What you did above was perfectly fine. Here's another way to do it with pipes (I'm just going to show one example here: 
(ACTUALLY - This is  a really hard problem to solve succinctly, so your way above is better!)

```{{r}}
# My lousy attempt
xtab_vars <- names(diabetes)[2:15]

for (i in 1:length(xtab_vars)) { 
  x <- 
  diabetes %>%
  group_by(CLASS, eval(parse(text = diabetes[,..i]))) %>%
  tally() %>%
  pivot_wider(names_from = CLASS,
              values_from = n)
  print(x)
  }
```

```{{r}}
#Logistic Regression Variable Selection -- Check for Correlations (p-value < 0.05)
cor.test(diabetes_data$Age, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Gender, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Polyuria, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Polydipsia, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$`sudden weight loss`, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$weakness, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Polyphagia, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$`Genital thrush`, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$`visual blurring`, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Itching, diabetes_data$class, method = "spearman") #diabetes_data$Itching <- NA
cor.test(diabetes_data$Irritability, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$`delayed healing`, diabetes_data$class, method = "spearman") #diabetes_data$'delayed healing' <- NA
cor.test(diabetes_data$`partial paresis`, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$`muscle stiffness`, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Alopecia, diabetes_data$class, method = "spearman")
cor.test(diabetes_data$Obesity, diabetes_data$class, method = "spearman") #diabetes_data$Obesity <- NA
```

We can accomplish the above with the following below more succinctly (and get a pretty picture to boot!) 

```{r}
# REF: https://stackoverflow.com/questions/62356378/spearman-correlation-plot-in-corrplot
library(corrplot)

diabetes %>% 
  mutate(across(everything(), as.numeric)) %>%
  cor(.,method = "spearman") %>%
  corrplot(method = "color",
           type = "upper",
           addCoef.col = "black")
```
```{{r}}
#Logistic Regression
logistic_regression <- glm(diabetes_data$class ~ diabetes_data$Age + 
      diabetes_data$Gender + 
      diabetes_data$Polyuria + 
      diabetes_data$Polydipsia + 
      diabetes_data$`sudden weight loss` + 
      diabetes_data$weakness + 
      diabetes_data$Polyphagia + 
      diabetes_data$`Genital thrush` + 
      diabetes_data$`visual blurring` + 
      diabetes_data$Irritability + 
      diabetes_data$`partial paresis` + 
      diabetes_data$`muscle stiffness` + 
      diabetes_data$Alopecia, family = "binomial", data = diabetes_data)
```

Here's a shorthand way to accomplish the above: 

```{r}
log_reg <- glm(CLASS ~ ., family = "binomial", data = diabetes)
summary(log_reg)
```

Below I'm coercing the CLASS variable to a numeric inline - it's shorter :) 

```{r}
cor.test(x = as.numeric(diabetes$CLASS), y = log_reg$fitted.values, method = "spearman")
```

Below is similar to what you did, I'm just performing some data transformations "on the fly". 

```{r graph-log-reg}
diabetes %>%
  mutate(across(where(is.factor), as.numeric)) %>%
  mutate(across(2:17, ~ .x - 1)) %>% 
  ggplot() + 
  geom_point(aes(x = log_reg$fitted.values, y = CLASS)) + 
  geom_smooth(aes(x = log_reg$fitted.values, y = CLASS),
              method = "glm",
              method.args = list(family = "binomial"),
              se = TRUE,
              show.legend = TRUE)
```

